---
title: "browsertime: cold page load, Chrome 73 vs Fenix 68 ARM on the Pixel 2"
author: Nick Alexander
date: April 12, 2019
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    includes:
      in_header: "html/header.html"
---

<style>
select {
    width: 30%;
    right: 10px;
    height: 30px;
}
</style>

```{r setup, include=FALSE}
library(glue)
library(dplyr)
library(DT)
library(ggplot2)
library(stringr)
library(tidyr)

btr <- read.csv('data/data.csv')
by_site <- btr %>%
  filter(proxy == 'replay') %>%
  select(-proxy, -timestamp)

by_site <- by_site %>% mutate(vehicle = case_when(
  grepl("Chrome/73.0", engine, fixed=TRUE) ~ "Chrome 73",
  grepl("Gecko/68.0", engine, fixed=TRUE) ~ "Fenix 68 ARM"
))

# Ordering the vehicle factor in this way makes the dot-plot below generally "follow reality": Fennec is slowest, then Fenix, then GVE, then Chrome.
by_site$vehicle <- factor(by_site$vehicle, levels=c("Fenix 68 ARM", "Chrome 73"))

by_site <- select(by_site, -engine)

runs <- data.frame(run = c(1),
                   device = factor(c("Pixel 2")),
                   iterations = c(9),
                   min_iterations = c(8))

with_min_iterations <- by_site %>%
  group_by(run, device, site, vehicle) %>%
  summarise(num_iterations=n()) %>%
  ungroup() %>%
  group_by(run, device, site) %>%
  # This ensures that every "vehicle" has 0 if there were no runs recorded for that vehicle configuration.
  spread(vehicle, num_iterations, fill = 0) %>%
  gather("vehicle", num_iterations, levels(by_site$vehicle)) %>%
  ungroup() %>%
  group_by(run, device, site) %>%
  summarise(min_iterations=min(num_iterations)) %>%
  ungroup()

# Each (run, device, site) that has sufficient data to draw conclusions.
sufficient <- left_join(with_min_iterations, runs %>% select(-iterations), by = c("run", "device")) %>%
  filter(min_iterations.x >= min_iterations.y) %>%
  select(run, device, site)

# The measurements corresponding to (run, device, site) tuples with sufficient data.
complete <- semi_join(by_site, sufficient, by = c("run", "device", "site"))

complete_sites <- complete %>% select(site) %>% distinct()
```

## Per-device, per-site vehicle comparisons

The following graph gives insight into how the vehicles compare on the
sites in the test corpus.  Here we see Chrome 73 and Fenix 68 ARM
generally competing well, with Chrome 73 having a slight edge.

The web.de URL appears to be an outlier: perhaps a bad recording
explains the aberrant measurements.

```{r, echo=FALSE, fig.width = 12, fig.height=12}
xx <- complete %>% group_by(device, run, site) %>% summarize(m=mean(pageLoadTime)) %>% ungroup() %>% arrange(m) %>% inner_join(complete, by=c('device', 'run', 'site'))
xx$site <- sub('https://', '', xx$site)
xx$site <- sub('http://', '', xx$site)
xx$site <- xx$site %>% str_trunc(30) %>% str_pad(30, 'right')

p <- ggplot(xx, aes(x=reorder(site, m), y=pageLoadTime))
p <- p + geom_point(aes(color=vehicle))
p <- p + facet_wrap(vars(device, run), ncol=1, scales='free', labeller=labeller(.multi_line=FALSE)) + theme(axis.text.x=element_text(angle=-67.5, hjust=0, vjust=0)) + xlab('site')
p <- p + theme(legend.justification = c(0, 1), legend.position = c(0.04, 0.995))
p
```

If we remove web.de and plot the median page load times, we can clearly see that Fenix 68 ARM is mostly lagging Chrome 73.

```{r, echo=FALSE, fig.width = 12, fig.height=12}
xx <- complete %>% filter(site != 'https://web.de')  %>% group_by(device, run, site) %>% summarize(m=mean(pageLoadTime)) %>% ungroup() %>% arrange(m) %>% inner_join(complete, by=c('device', 'run', 'site'))
xx$site <- sub('https://', '', xx$site)
xx$site <- sub('http://', '', xx$site)
xx$site <- xx$site %>% str_trunc(30) %>% str_pad(30, 'right')

xx <- xx %>%
  filter(!grepl('web.de', site, fixed=TRUE)) %>%
  group_by(device, run, site, vehicle) %>%
  summarize(m=first(m), med=median(pageLoadTime)) %>%
  ungroup()

# xx$scaled_med <- xx$med * 1.1
# xx$scaled_med[xx$vehicle != 'Chrome 73'] <- NA

p <- ggplot(xx, aes(x=reorder(site, m), y=med, fill=vehicle))
# p <- p + geom_errorbar(aes(ymin = 0, ymax = scaled_med), width=0.6)
p <- p + geom_bar(position="dodge", stat="identity")
p <- p + facet_wrap(vars(device, run), ncol=1, scales='free', labeller=labeller(.multi_line=FALSE)) + theme(axis.text.x=element_text(angle=-67.5, hjust=0, vjust=0)) + xlab('site')
p <- p + theme(legend.justification = c(0, 1), legend.position = c(0.04, 0.995))
p
```

## Gecko engine vehicles over Chrome 73

```{r g_over_c, include=FALSE}
g_over_c <- complete %>%
  group_by(run, device, site, vehicle) %>%
  summarise(pageLoadTime_mean=mean(pageLoadTime)) %>% # , pageLoadTime_rsd=sd(pageLoadTime)/mean(pageLoadTime)*100) %>%
  ungroup() %>%
  group_by(run, device, site) %>%
  spread(vehicle, pageLoadTime_mean) %>%
  ungroup()

g_over_c[c("Fenix 68 ARM")] <- g_over_c[c("Fenix 68 ARM")] / g_over_c$`Chrome 73`

g_over_c <- g_over_c %>% select(-`Chrome 73`, -run)
```

```{r echo=FALSE}
datatable(g_over_c, options=list(dom='t', pageLength = nrow(g_over_c)),
          fillContainer = FALSE,
          caption=('Gecko engine vehicles pageload over Chrome pageload (lower is better)' %>% glue())) %>%
  # formatRound(c(3, 4), 2) %>%
  formatPercentage(c("Fenix 68 ARM"))
```
